On the specific recursive structure of iterated parametricity

### 2025-11-09 paper/paper.tex
A couple of fixes/additions

    \framep[n][0]                                                                          &

@@

\begin{align*}
  \depstype[n][0,..,p-1] \defeq
  \left(\begin{array}{l}
            \framep[n][0,...,p-1] :\frametype[n][0,...,p-1] \\
            \painting[n][0,...,p-1] :\paintingtype[n][0,...,p-1]
            (\framep[n][0,...,p-1])                         \\
            \restrf[n][0,...,p-1]  :\restrftype[n][0,...,p-1]
            \left(\begin{array}{l}
                \framep[n][0,...,p-1]   \\
                \painting[n][0,...,p-1] \\
              \end{array}\right)                   \\
          \end{array}\right)
\end{align*}
so that the type of $\framep[n][0,...,p]$ can be shortened into $\depstype[n][0,..,p-1] \rightarrow \frametype[n][0,...,p]$ and similarly for $\layertype[n][p]$ and $\layer[n][p]$.
Later, we shall write $\deps.\framep$, $\deps.\painting$, $\deps.\restrf$ for the corresponding projections of a $\deps$. By abuse of notation, we also write $\unitpoint$ for the only inhabitant of $\depstype[n][0,..,0-1]$.

@@

Let us introduce another abbreviation to combine a full $\deps[n][0,n]$ with a family $E:\framep[n-1][n-1]\rightarrow \U$:
\begin{align*}
  \fulldepstype[n] \defeq
  \left(\begin{array}{l}
            \deps[n][0,...,n-1] :\depstype[n][0,...,n-1] \\
            E:\frametype[n-1][n-1]\rightarrow \U         \\
          \end{array}\right)
\end{align*}
TODO: say the convention on projection in the Introduction

@@

We can now formally define $\painting[n][p]$ by recursion on $n-p$:
\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \painting[n][p]      &
    (\fulldeps[n-1]:\fulldepstype[n-1])           \\
    \painting[n][n]      &
    (\fulldeps[n-1])     & \defeq & \fulldeps.E   \\
    \painting[n][p](p<n) &
    (\fulldeps[n-1])     & \defeq & d \mapsto
    \left(\begin{array}{l}
              \Sigma l:\layer(\deps[n-1][0,...,p-1]). \\
              \painting[n][p+1,...,n-1]
              (\fulldeps[n-1])(d,l)                   \\
            \end{array}\right) \\
  \end{array}
\end{equation*}

@@

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \painting[n][0,...,p]   &
    (\fulldeps[n-1])        & :      & \paintingtype[n][0,...,p](\framep[n][0,...,p](\fulldeps.\deps[n-1][0,...,p])) \\
    \painting[n][0,...,0-1] &
    \left(\begin{array}{l}
              \unitpoint \\
              E          \\
            \end{array}\right)  & \defeq & \unittype                                                                   \\
    \painting[n][0,...,p]   &
    (\fulldeps[n-1])
                            & \defeq &
    (\painting[n][0,...,p-1](\fulldeps[n-1]),\painting[n][p](\fulldeps[n-1]))                                        \\
  \end{array}
\end{equation*}

At this stage, we are able to define $\framep$, $\painting$ and $\restrftype$ from $\deps$ at the respective previous level. To go further, we need to define the sequence $\restrptype$ of types of painting restrictions:

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \restrptype[n][0,...,p]             &
    (\fulldeps[n-1]:\fulldepstype[n-1]) & :          & \U                                                    \\
    \restrptype[n][]                    & \unitpoint & \defeq                                    & \unittype \\
    \restrptype[n][0,...,p]             &
    (\fulldeps[n-1])
                                        & \defeq     & \restrptype[n+1][0,...,p-1] \times \left(
    \begin{array}{l}
        \Pi q\leq n.\Pi \epsilon.\Pi d:\framep(\deps[n][0,...,p]). \\
        \painting[n][p]
        (\fulldeps[n-1])(d) \rightarrow                            \\ \fulldeps.\deps.\painting[n][p](\fulldeps.\deps.\restrf[n][p][q][\epsilon](d))\end{array}
    \right)                                                                                                  \\
  \end{array}
\end{equation*}

Thanks to $\restrptype$, we can mutually define $\restrf$ and $\cohftype$, using abbreviations in passing for $\layer$ and $\layertype$:

@@

                \fulldeps[n]:\depstype[n]   \\

@@

                \fulldeps.\deps.\framep[n][0,...,p]   \\
                \fulldeps.\deps.\painting[n][0,...,p] \\

@@

      \deps.\restrf[n][p][q][\epsilon](\restrf[n+1][p][r][\omega]

@@

      \deps.\restrf[n][p][r][\omega](\restrf[n+1][p][q+1][\epsilon]

@@

@@

            \painting[n][p](\fulldeps.\deps[n][0,...,n]) \\

@@

                      \fulldeps.\deps[n][0,...,p] \\

@@

        \fulldeps.\deps.\layer[n][p](\deps[n][0,...,n])(\restrf[n][p]

@@

    \lambda q\leq n. \lambda \epsilon.\lambda\omega.\,\overrightarrow{\cohf[n][p][q][q][\epsilon][\omega]}{\restrp[n][p](l \omega)}                          \\

@@

\begin{align*}
  \depscohstype[n][0,..,p-1] \defeq
  \left(\begin{array}{l}
            \deps[n][0,...,n] :\depstype[n][0,...,n]        \\
            \restrp[n][0,...p-1]: \restrptype[n][0,...,p-1] \\
            \cohf[n][0,...,p-1]: \cohftype[n][0,...,p-1]    \\
          \end{array}\right)
\end{align*}
so that the type of $\restrf[n][0,...,p]$ can be rephrased as $\Pi \depscohs[n][0,...,p-1]:\depscohstype[n][0,..,p-1]. \restrftype[n][0,...,p]      \left(\begin{array}{l}
      \depscohs.\fulldeps.\deps.\framep[n][0,...,p]   \\
      \depscohs.\fulldeps.\deps.\painting[n][0,...,p] \\
    \end{array}\right)$, and similarly for $\restrltype$ and $\restrl$. Later, we shall write $\depscohs.\deps$, $\depscohs.\restrp$, and $\depscohs.\cohf$ for the corresponding projections of an element $\depscohs$ in $\depscohstype$. By abuse of notation, we also write $\unitpoint$ for the only inhabitant of $\depscohstype[n][0,..,0-1]$.

@@

\begin{align*}
  \overline{\deps}(\depscohs:\depscohstype) & :      & \depstype \\
  \overline{\deps}(\depscohs)               & \defeq &
  \left(\begin{array}{l}
            \framep[n][p](\depscohs.\deps[n][0,...,p])   \\
            \painting[n][p](\depscohs.\deps[n][0,...,n]) \\
            \restrf[n][p](\depscohs[n][0,...,p])         \\
          \end{array}\right)             \\
\end{align*}

We need also the notion of full dependencies at the level of coherence conditions, that combines a full $\depscohs[n][0,...,n]$ with a set $E:\framep[n][n](\overline{\deps}(\depscohs[n][0,...,n-1])) \rightarrow \U$:
\begin{align*}
  \fulldepscohstype[n] \defeq
  \left(\begin{array}{l}
            \depscohs[n][0,...,n-1] :\depscohstype[n][0,...,n-1]                         \\
            E:\framep[n-1][n-1](\overline{\deps}(\depscohs[n][0,...,n-1]))\rightarrow \U \\
          \end{array}\right)
\end{align*}
TODO: say the convention on projection in the Introduction

@@

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \restrp[n][p]                           &
    (\fulldepscohs[n]:\fulldepscohstype[n]) & :      &
    \left(
    \begin{array}{l}
        \Pi q\leq n. \Pi \epsilon. \Pi d:\framep[n][p]
        \left(\begin{array}{l}
                  \overline{\deps}(\depscohs)    \\
                  E':\framep[n][n]\rightarrow \U \\
                \end{array}\right).    \\
        \painting[n][p]
        \left(\begin{array}{l}
                  \overline{\deps}(\depscohs)    \\
                  E':\framep[n][n]\rightarrow \U \\
                \end{array}\right)(d) \rightarrow \\
        \painting[n-1][p]
        \left(\begin{array}{l}
                  \depscohs[n][0,...,n].\deps       \\
                  E:\framep[n-1][n-1]\rightarrow \U \\
                \end{array}\right)(\restrf[n][p](\depscohs[n][0,...,n].deps)(d))\end{array}\right)                                       \\
    \restrp[n][n]                           &
    (\fulldepscohs[n][0,...,n])             & \defeq & \lambda q\leq n.\lambda \epsilon. \lambda d. \lambda (l,c). l\epsilon \\
    \restrp[n][p](p<n)                      &
    (\fulldepscohs[n][0,...,n])             & \defeq & \lambda q\leq n.\lambda \epsilon. \lambda d. \lambda (l,c).
    \left(\begin{array}{l}
              \restrl[n][p][q][\epsilon](\fulldepscohs.\depscohs[n-1][0,...,p-1])(l),               \\
              \restrp[n][p][q][\epsilon]
              (\fulldepscohs[n][0,...,n])(\restrf[n][p](\fulldepscohs.\depscohs[n][0,...,p])(d))(c) \\
            \end{array}\right)                              \\
  \end{array}
\end{equation*}

@@

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \cohptype[n][0,...,p] & (\depscohs[n][0,...,n]:\depscohstype[n][0,...,n]) & :      & \U                                      \\
    \cohptype[n][]        & \unitpoint                                        & \defeq & \unittype                               \\
    \cohptype[n][0,...,p] & (\depscohs[n][0,...,n])                           & \defeq & \cohptype[n+1][0,...,p-1] \times \left(
    \begin{array}{l}
        \Pi q\leq n.\,\Pi r\leq q.\,\Pi \epsilon.\,\Pi \omega. \\
        \depscohs.\restrp[n][p][q][\epsilon](\restrp[n+1][p][r][\omega]
        \left(\begin{array}{l}
                  \depscohs[n+1][0,...,n] \\
                  \cohp[n+1][0,...,p-1]   \\
                \end{array}\right)(d)) =                         \\
        \depscohs.\restrp[n][p][r][\omega](\restrp[n+1][p][q+1][\epsilon]
        \left(\begin{array}{l}
                  \depscohs[n+1][0,...,n] \\
                  \cohp[n+1][0,...,p-1]   \\
                \end{array}\right)(d))
      \end{array}
    \right)                                                                                                                      \\
  \end{array}
\end{equation*}

>>>>

    \framep[n][0,]                                                                          &

@@

\begin{align*}
  \depstype[n][0,..,p-1] \defeq
  \left(\begin{array}{lll}
            \framep[n][0,...,p-1] &:&\frametype[n][0,...,p-1] \\
            \painting[n][0,...,p-1] &:&\paintingtype[n][0,...,p-1]
            (\framep[n][0,...,p-1])                         \\
            \restrf[n][0,...,p-1]  &:&\restrftype[n][0,...,p-1]
            \left(\begin{array}{l}
                \framep[n][0,...,p-1]   \\
                \painting[n][0,...,p-1] \\
              \end{array}\right)                   \\
          \end{array}\right)
\end{align*}
so that the type of $\framep[n][0,...,p]$ can be shortened into $\depstype[n][0,..,p-1] \rightarrow \frametype[n][0,...,p]$ and similarly for $\layertype[n][p]$ and $\layer[n][p]$.
Later, we shall write $\deps.\framep$, $\deps.\painting$, $\deps.\restrf$ for the corresponding projections of a $\deps$. By abuse of notation, we also write $\unitpoint$ for the only inhabitant of $\depstype[n][0,..,0-1]$.

@@

Let us introduce another abbreviation to combine a full $\deps[n][0,n]$ with a family $E:\framep[n-1][n-1]\rightarrow \U$:
\begin{align*}
  \fulldepstype[n] \defeq
  \left(\begin{array}{lll}
            \deps[n][0,...,n-1] &:&\depstype[n][0,...,n-1] \\
            E&:&\frame[n-1][n-1](\deps[n][0,...,n-1]) \rightarrow \U         \\
          \end{array}\right)
\end{align*}
TODO: say the convention on projection in the Introduction

@@

We can now formally define $\painting[n][p]$ by recursion on $n-p$:
\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \painting[n][p]      &
    (\fulldeps[n-1]:\fulldepstype[n-1])           \\
    \painting[n][n]      &
    (\fulldeps[n-1])     & \defeq & \fulldeps.E   \\
    \painting[n][p](p<n) &
    (\fulldeps[n-1])     & \defeq & d \mapsto
    \left(\begin{array}{l}
              \Sigma l:\layer(\fulldeps.\deps[n-1][0,...,p-1]). \\
              \painting[n][p+1,...,n-1]
              (\fulldeps[n-1])(d,l)                   \\
            \end{array}\right) \\
  \end{array}
\end{equation*}

@@

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \painting[n][0,...,p]   &
    (\fulldeps[n-1])        & :      & \paintingtype[n][0,...,p](\framep[n][0,...,p](\fulldeps.\deps[n-1][0,...,p])) \\
    \painting[n][0,...,0-1] &
    E  & \defeq & \unittype                                                                   \\
    \painting[n][0,...,p]   &
    (\fulldeps[n-1])
                            & \defeq &
    (\painting[n][0,...,p-1](\fulldeps[n-1]),\painting[n][p](\fulldeps[n-1]))                                        \\
  \end{array}
\end{equation*}

At this stage, we are able to define $\framep[n][0,...,n]$, $\painting[n][0,...,n]$ and $\restrftype[n][0,...,n-1]$ from $\framep[n-1][0,...,n-1]$, $\painting[n-1][0,...,n-1]$ and $\restrf[n-1][0,...,n-1]$, that is from $\deps[n-1][0,...,n-1]$. To go further, we need to define the sequence $\restrptype[n][0,...,p]$ of types of painting restrictions:

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \restrptype[n][0,...,p]             &
    (\fulldeps[n-1]:\fulldepstype[n-1]) & :          & \U                                                    \\
    \restrptype[n][]                    & E & \defeq                                    & \unittype \\
    \restrptype[n][0,...,p]             &
    (\fulldeps[n-1])
                                        & \defeq     & \restrptype[n+1][0,...,p-1] \times \left(
    \begin{array}{l}
        \Pi q\leq n.\Pi \epsilon.\Pi d:\framep[n][p](\fulldeps.\deps[n][0,...,p]). \\
        \painting[n][p]
        (\fulldeps[n-1])(d) \rightarrow                            \\ \fulldeps.\deps.\painting[n][p](\fulldeps.\deps.\restrf[n][p][q][\epsilon](d))\end{array}
    \right)                                                                                                  \\
  \end{array}
\end{equation*}

Thanks to $\restrptype[n][0,...,n]$, we can mutually define $\restrf[n][p]$ and $\cohftype[n][p]$ from ``full dependencies'', that is from $\frame[n][0,...,n]$, $\painting[n][0,...,n]$, $\restrf[n][0,...,n]$ and some $E:\frame[n+1][n+1](\, using abbreviations in passing for $\layer$ and $\layertype$:

@@

@@

                \framep[n][0,...,p](\fulldeps.\deps[n][0,...,p])   \\
                \painting[n][0,...,p](\fulldeps[n]) \\

@@

      \Pi q\leq n.\,\Pi r\leq q.\,\Pi \epsilon.\,\Pi \omega.\,\Pi d:\framep[n][p]
                \frame(\fulldeps.\deps[n][0,...,p])          \\
                \painting[n+1][0,...,p-1](\fulldeps[n]) \\
                \restrf[n][0,...,p]

@@

      \\
      \fulldeps.\deps.\restrf[n][p][q][\epsilon](\restrf[n+1][p][r][\omega]

@@

      \fulldeps.\deps.\restrf[n][p][r][\omega](\restrf[n+1][p][q+1][\epsilon]

@@

            \painting[n][p](\fulldeps[n]) \\

@@

@@

        \layer[n][p](\fulldeps.\deps[n][0,...,n])(\restrf[n][p]

@@

    \lambda q\leq n. \lambda \epsilon.\,\lambda d.\lambda l.\overrightarrow{\cohf[n][p][q][q][\epsilon][\omega](d)}{\restrp[n][p][q][\epsilon](\restrf[n][p][q][\epsilon])(l \omega)}                          \\

@@

\begin{align*}
  \depscohstype[n][0,..,p-1] \defeq
  \left(\begin{array}{l}
            \fulldeps[n][0,...,n] &:&\fulldepstype[n][0,...,n]        \\
            \restrp[n][0,...p-1]&:& \restrptype[n][0,...,p-1] \\
            \cohf[n][0,...,p-1]&:& \cohftype[n][0,...,p-1]    \\
          \end{array}\right)
\end{align*}
so that the type of $\restrf[n][0,...,p]$ can be rephrased as $\Pi \depscohs[n][0,...,p-1]:\depscohstype[n][0,..,p-1]. \restrftype[n][0,...,p]      \left(\begin{array}{l}
      \depscohs.\fulldeps.\deps.\framep[n][0,...,p]   \\
      \depscohs.\fulldeps.\deps.\painting[n][0,...,p] \\
    \end{array}\right)$, and similarly for $\restrltype$ and $\restrl$. Later, we shall write $\depscohs.\deps$, $\depscohs.\restrp$, and $\depscohs.\cohf$ for the corresponding projections of an element $\depscohs$ in $\depscohstype$. By abuse of notation, we also write $\unitpoint$ for the only inhabitant of $\depscohstype[n][0,..,0-1]$.

@@

\begin{align*}
  \overline{\deps}(\depscohs:\depscohstype) & :      & \depstype \\
  \overline{\deps}(\depscohs)               & \defeq &
  \left(\begin{array}{l}
            \framep[n][p](\depscohs.\fulldeps.\deps[n][0,...,p])   \\
            \painting[n][p](\depscohs.\fulldeps[n][0,...,n]) \\
            \restrf[n][p](\depscohs[n][0,...,p])         \\
          \end{array}\right)             \\
\end{align*}

\begin{align*}
  \overline{\fulldeps}(\fulldepscohs:\fulldepscohstype) & :      & \fulldepstype \\
  \overline{\fulldeps}(\fulldepscohs)               & \defeq &
  \left(\begin{array}{l}
            \overline{\deps}(\depscohs)\\
            \depscohs.\fulldeps.E\\
          \end{array}\right)             \\
\end{align*}

We need also the notion of full dependencies at the level of coherence conditions, that combines a full $\depscohs[n][0,...,n]$ with a set $E:\framep[n][n](\overline{\deps}(\depscohs[n][0,...,n-1])) \rightarrow \U$:
\begin{align*}
  \fulldepscohstype[n] \defeq
  \left(\begin{array}{l}
            \depscohs[n][0,...,n-1] &:&\depscohstype[n][0,...,n-1]                         \\
            E&:&\framep[n-1][n-1](\overline{\deps}(\depscohs[n][0,...,n-1]))\rightarrow \U \\
          \end{array}\right)
\end{align*}
TODO: say the convention on projection in the Introduction

@@

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \restrp[n][p]                           &
    (\fulldepscohs[n]:\fulldepscohstype[n]) & :      &
    \left(
    \begin{array}{l}
        \Pi q\leq n. \Pi \epsilon. \Pi d:\framep[n][p]
        (\overline{\deps}(\depscohs)).    \\
        \painting[n][p]
        (\overline{\fulldeps}(\fulldepscohs[n])(d) \rightarrow \\
        \painting[n-1][p](\fulldepscohs.\depscohs.\fulldeps[n])
          (\restrf[n][p](\fulldepscohs.\depscohs[n][0,...,n])(d))\end{array}\right)                                       \\
    \restrp[n][n]                           &
    (\fulldepscohs[n][0,...,n])             & \defeq & \lambda q\leq n.\lambda \epsilon. \lambda d. \lambda (l,c). l\epsilon \\
    \restrp[n][p](p<n)                      &
    (\fulldepscohs[n][0,...,n])             & \defeq & \lambda q\leq n.\lambda \epsilon. \lambda d. \lambda (l,c).
    \left(\begin{array}{l}
              \restrl[n][p][q][\epsilon](\fulldepscohs.\depscohs[n-1][0,...,p-1])(l),               \\
              \restrp[n][p][q][\epsilon]
              (\fulldepscohs[n])(\restrf[n][p](\fulldepscohs.\depscohs[n][0,...,p])(d))(c) \\
            \end{array}\right)                              \\
  \end{array}
\end{equation*}

@@

\begin{equation*}
  \fontsize{7.2}{9}\selectfont
  \begin{array}{llcl}
    \cohptype[n][0,...,p] & (\fulldepscohs[n]:\fulldepscohstype[n]) & :      & \U                                      \\
    \cohptype[n][]        & \unitpoint                                        & \defeq & \unittype                               \\
    \cohptype[n][0,...,p] & (\fulldepscohs[n])                           & \defeq & \cohptype[n+1][0,...,p-1] \times \left(
    \begin{array}{l}
        \Pi q\leq n.\,\Pi r\leq q.\,\Pi \epsilon.\,\Pi \omega.\,\Pi d:...\,\Pi c \\
        \fulldepscohs.\depscohs.\restrp[n][p][q][\epsilon](\restrf(d))(\restrp[n+1][p][r][\omega]
        (\fulldepscohs[n+1])(d)(c)) =                         \\
        \fulldepscohs.\depscohs.\restrp[n][p][r][\omega](\restrf(d))(\restrp[n+1][p][q+1][\epsilon]
        (\fulldepscohs[n+1])(d)(c))
      \end{array}
    \right)                                                                                                                      \\
  \end{array}
\end{equation*}
